\documentclass[a4paper,12pt]{book}
\usepackage{graphicx} % For graphics
\usepackage{float}
\usepackage{tikz}
\usepackage{tkz-euclide} % Add this line to include the tkz-euclide package
\usetikzlibrary{calc}
\usepackage{amsmath}
\usepackage{enumitem}
\usepackage{amssymb}
\usepackage{pgfplots}
\pgfplotsset{compat=newest}
\usepgfplotslibrary{fillbetween} 

% Configure hyperref package to color links blue
\usepackage[colorlinks=true, linkcolor=blue, citecolor=blue, urlcolor=blue]{hyperref}

\newcommand{\inputtoc}[1]{\input{#1}}

% \makeindex % Command to make the index

\pgfmathdeclarefunction{csc}{1}{%
  \pgfmathparse{1/sin(#1)}%
}

\newenvironment{exercise}[1][]
  {\par\medskip\noindent\textbf{Exercise #1} \rmfamily}
  {\medskip}

% Define the problem environment
\newcounter{problem}
\newenvironment{problem}[1][\theproblem]
{\refstepcounter{problem}\par\medskip\noindent\textbf{Problem~#1:} \rmfamily}{\medskip}

% Define the solution environment
\newenvironment{solution}[1][]
{\par\noindent\textit{Solution:} \rmfamily}{\medskip}

% Define the example environment
\newcounter{example}
\newenvironment{example}[1][\theexample]
  {\refstepcounter{example}\par\medskip\noindent\textbf{Example~#1:} \rmfamily}
  {\medskip}
  
% Define theorem, application, definition, method, and discussion environments
\newtheorem{theorem}{Theorem}
\newtheorem{application}{Application}
\newtheorem{definition}{Definition}
\newtheorem{method}{Method}
\newtheorem{discussion}{Discussion}

\begin{document}

% --- Book Cover ---
\begin{titlepage}
    \begin{tikzpicture}[remember picture, overlay]
        % Background color
        \fill[cyan!30] (current page.south west) rectangle (current page.north east);

        % Decorative Circles Pattern at the bottom
        \foreach \i in {0,...,36}
        {
            \fill[cyan!40, rotate around={10*\i:(current page.center)}] 
                (current page.south) circle (1cm);
        }
        
        % Title and additional info
        \node at (current page.center) [font=\Huge, text width=0.8\textwidth, align=center] 
            {\textbf{Calculus in Brief with Python, Colab, GitHub, and \LaTeX  \\ CiB - Version 0.41}};
            
        \node[align=center, font=\large] at (current page.center) 
            [yshift=-2.5cm] {\today}; % <- This is the compilation date
            
        \node[align=center, font=\large] at (current page.center) 
            [yshift=-3.5cm] {MIT License};
        
        \node[align=center, font=\large] at (current page.center) 
            [yshift=-5cm] {Available on GitHub at: \\ 
            \url{https://GitHub.com/nicholaskarlson/CiB}};
    \end{tikzpicture}
\end{titlepage}

% --- Table of Contents ---
\tableofcontents
\cleardoublepage

% --- Preface ---
\chapter*{Preface}
\addcontentsline{toc}{chapter}{Preface}
This text, \emph{Calculus in Brief - CiB} aspires to be more than just another math book. This book strives to foster collaborative math writing. Note that this book has very few references. The reader is encouraged to use resources available on the Web to fact-check. This book's view on ``causation'' and facts is heavily influenced by Mosteller and Tukey \cite{mosteller1977}.

\section*{Redefining the Role of the Reader}
Calculus in Brief (CiB) is an endeavor to reshape how math is written, understood, and studied. It's not just a passive read but an open-source approach to math, aiming to encourage students to become proactive learners.

This project strives to break the traditional mold of math education and invites readers and professional mathematicians to participate actively.

\section*{A Dynamic Relationship with Math}
\emph{Calculus in Brief} is not just a book but a movement and methodology, heralding a new era in how we approach, consume, and interact with math. By positioning the reader as an integral part of the math-book process, CiB fosters a dynamic relationship with math, making mathematics more accessible, proactive, and relevant. In this shifting paradigm, we are all potential mathematicians, creators of interesting and relevant ways to learn and study math.

Please fork the LaTeX source code for CiB (available on GitHub) and create your own book that chooses the facts and exercises most relevant to you! Also, starring the CiB project on GitHub would be greatly appreciated! Thanks for reading CiB!

% --- Chapters ---
\chapter{Introduction to CiB}
\subsection*{Welcome to CiB on GitHub}
Calculus in Brief, abbreviated CiB, isn't merely a passive read. It's an endeavor to reshape how math is written, studied, and taught. By presenting an open-source approach to math, the goal is to encourage everyone to become proactive readers and writers of math. 

\subsection*{Fostering a Proactive Engagement with Math}

\emph{Calculus in Brief} is a call for a renewed engagement with mathematics. CiB is an endeavor to reshape how math is written, understood, and studied. It's not just a passive read but an open-source approach to math, aiming to encourage students to become proactive learners.

This project strives to break the traditional mold of math education and encourages readers and professional mathematicians to participate actively.

\bigskip
\noindent
Please fork the \LaTeX{} source code for CiB (available on GitHub) and create your own book on Calculus that chooses the content most relevant to you! Also, starring the CiB project on GitHub would be greatly appreciated! Thanks for reading CiB!

\chapter{Open-Source Ethos}
\section*{The Spirit of Shared Knowledge and Collaboration}
Math, like software, is better when it's open. CiB draws inspiration from the open-source software movement; this section elucidates how a collaborative, transparent, and shared approach can enhance our understanding of math. Here, we look at the philosophy behind open-source and how it beautifully combines with the study of mathematics.

\subsection*{Open-Source Math: Preserving Tradition Through Collaborative Exploration}
Mathematics, like software, thrives when it embraces openness and transparency. CiB takes a leaf from the proven benefits of the open-source software model; this section highlights how a collaborative and transparent method can improve and deepen our grasp of math and its texts. Here, we explore the principles of open source and how these principles align with the development of mathematics and its texts.

\subsection*{Understanding the Open-Source Ethos}
The open-source paradigm revolves around shared ownership, collaboration, and the free exchange of knowledge. In the software realm, this approach has led to groundbreaking innovations built and enhanced by a global community of skilled contributors. United by a mutual objective, these individuals pool their diverse talents and insights to improve and share software solutions for broader public benefit.

\subsection*{Advantages of the Open-Source Framework in Math}
\subsubsection*{Collective Insight}
Mirroring the collaborative essence of open-source software, many individuals can offer their perspectives and knowledge, making math texts more robust and varied.

\subsubsection*{Enhancement and Accuracy}
Open platforms foster an environment of constructive criticism, ensuring prompt identification and correction of inaccuracies. This meticulous peer review can help provide a credible and current mathematical text.

\subsubsection*{Universal Access}
Much as open-source software promotes free access and modification, open-source math prioritizes universal accessibility. This ensures mathematics knowledge isn't restricted to a select few but is available to all curious minds.

\subsection*{Potential Challenges}
Despite its advantages, melding open-source with math has potential pitfalls. The volume of contributions can complicate accuracy verification processes. 

However, the very community championing this open-source approach to math can serve as its vigilant protector. They can ensure that contributions undergo rigorous evaluation and referencing, akin to the meticulous checks within the open-source software community.

\subsection*{Conclusion: Reinvigorating Our Experience with Math}
Adopting an open-source perspective to the approach of math signifies a refreshed approach. It beckons a worldwide community to collaborate and forge a comprehensive and exciting math text. In this refreshed approach, every individual can play a part, both as a contributor and a learner. Math texts, through this lens, evolve and flourish, reflecting the collective input of active participants.

\chapter{Introduction to GitHub}
\section*{The Hub for Modern Collaboration}
\subsection*{Harnessing GitHub: A New Frontier in Collaborative Math Writing}
At the heart of our collaborative math endeavor lies GitHub, a platform traditionally associated with code but now repurposed for our endeavor. This section provides a primer on GitHub, laying the foundation for those unfamiliar and offering insights into its transformative potential for collective math writing, learning, and teaching.

\subsection*{A Brief Introduction to GitHub}
Originally conceptualized as a platform for developers, GitHub is a repository hosting service that facilitates version control using Git. At its core, it allows multiple users to work on a project simultaneously, tracking changes and ensuring that the latest version of a project is always accessible. Over the years, GitHub has grown beyond its initial software-centric confines, becoming a hub for all kinds of collaborative projects, from writing to data science and now to math.

\subsection*{Repurposing GitHub for Math Texts}
\subsubsection*{Version Control}
Math writing, like software, is dynamic and constantly evolving. As new sources or perspectives emerge, math texts may need revisions. GitHub's version control ensures that every change made to a document is tracked, enabling mathematicians to see how math texts evolve over time.

\subsubsection*{Collaborative Writing}
Multiple contributors can work on a single math text simultaneously. This multi-user capability ensures diverse viewpoints can be seamlessly integrated, making the math text richer and more comprehensive.

\subsubsection*{Review and Feedback}
Just as developers review and comment on code, mathematicians can provide feedback on written content. This feature encourages rigorous peer review, ensuring accuracy and credibility.

\subsubsection*{Open Access}
Math texts on GitHub can be made public, granting anyone access to read, contribute, or fork the text into their own versions. This workflow democratizes math texts, making the creation process a collective endeavor rather than the domain of a select few.

\subsubsection*{Transparency}
All changes and contributions are logged, providing a clear trail of the evolution of a mathematical text. This transparency bolsters the credibility of the text hosted on the platform.

\subsubsection*{Community Building}
Beyond just writing, GitHub fosters a community of mathematicians, enthusiasts, and readers who can discuss, debate, and engage in meaningful dialogues about math and available math texts on GitHub.

\subsection*{Conclusion: Envisioning a Collaborative Mathematical Landscape}
Embracing GitHub as a tool for collaborative math signifies more than just a shift in approach; it heralds a new era of inclusivity, transparency, and dynamism in writing, learning, and math teaching. 

\chapter{Encouragement to Fork}
\subsection*{Invitation to Dive Deep and Make It Your Own}
CiB isn't a static entity. It thrives on evolution, adaptation, and diversification, much like math itself. We encourage readers to "fork" (a term soon to be discussed) and create their own versions of this book. Read this section to understand the essence of "forking" and how it can be the starting point of your unique math journey.

\subsection*{The Concept of Forking: A Brief Overview}

In the realm of software development, particularly in platforms like GitHub, "forking" refers to the act of creating a copy of a project, allowing one to make changes independently of the original. In this context, forking CiB enables readers to take the base content and adapt, modify, and expand upon it, tailoring the narrative to resonate with their perspectives, insights, and understanding.

\subsection*{How to Begin Your Forking Journey}

Start Small: You don't need to rewrite entire chapters. Begin by adding annotations, insights, or even footnotes to existing content. As you grow more confident, you can expand and modify larger sections.

Engage with the Community: Share your forked version with other readers. This encourages discourse, debate, and constructive feedback, allowing your text to be refined and enhanced.

Celebrate input: Encourage others around you to fork and create their own versions. The more in-depth the input, the deeper our collective understanding of math potentially becomes.

\subsection*{Conclusion: The Power of Collective Math}

The invitation to fork CiB isn't just about creating different versions of a book. It's a call to embrace collective writing, learning, and teaching. By embracing the essence of forking, math is not just something we read but something we actively shape, share, and pass on.

\chapter{More About GitHub}
\section*{Discovering the Power of Collaborative Tools}
Diving deeper into the world of GitHub, this chapter provides a comprehensive overview. Beyond its technicalities, we explore how GitHub emerged as a revolutionary platform for collaboration and how it can be leveraged for those interested in writing, teaching, and learning about math.

\subsection*{The Genesis of GitHub}
GitHub began as a platform designed for software developers to manage and track changes to their codebase. Launched in 2008, it swiftly gained traction due to its user-friendly interface and efficient version control system powered by Git. Over the years, it evolved from a mere repository hosting service to a dynamic hub of collaboration, housing millions of projects and engaging tens of millions of users worldwide.

\subsection*{GitHub: More than Just Code}
While GitHub's origins are rooted in code collaboration, its adaptable nature has made it a favored platform for various non-code projects. Writers, designers, educators, and researchers have discovered the potential of GitHub as a tool for:

\subsubsection*{Document Collaboration}
With its built-in version control, contributors can track changes, revert to previous versions, and seamlessly merge updates.

\subsubsection*{Project Management}
With features like "issues" and "milestones," teams can organize tasks, set goals, and monitor progress.

\subsubsection*{Open Access \& Transparency}
Public repositories allow for open contributions, ensuring transparency and fostering a sense of collective ownership.

\subsubsection*{Collaborative Writing}
Multiple contributors can simultaneously work on a single document, with every change being tracked and attributed, facilitating teamwork on extensive projects like books or research papers.

\subsubsection*{Engaging the Public}
With the platform's inherent transparency, researchers can make their work-in-progress accessible to the public, inviting insights, corrections, and contributions.

\subsection*{Case Study: CiB's Use of GitHub}
CiB's journey on GitHub is a testament to the platform's potential in mathematical endeavors. By hosting the book on GitHub, the following is possible:

\subsubsection*{Feedback Loop}
Readers can raise "issues," pointing out inaccuracies, suggesting enhancements, or even recommending new sections or topics.

\subsubsection*{Forking}
As previously discussed, readers can "fork" the repository, creating their unique versions of the book while staying connected to the original.

\subsubsection*{Regular Updates}
With math being dynamic, the book can be regularly updated, with new versions being released as and when significant changes are incorporated.

\subsection*{Challenges and Considerations}
While GitHub offers many advantages, it's essential to understand its limitations:

\subsubsection*{Learning Curve}
For those unfamiliar with Git or version control, there can be an initial learning curve.

\subsubsection*{Data Overwhelm}
With vast amounts of data and contributions, ensuring quality and accuracy can be challenging.

\subsubsection*{Diverse Audience Management}
Catering to both tech-savvy and non-tech audiences might require creating additional resources or tutorials to ensure inclusivity.

\subsection*{Conclusion: GitHub – A Paradigm Shift in Collaboration}
The rise of GitHub marks a significant shift in how we perceive and participate in collaborative projects. Its adaptability, transparency, and user-centric design make it a powerful tool, not just for coders but for anyone passionate about collective endeavors. In the realm of mathematics, GitHub promises a future where texts are continually refined, expanded, and enriched by a global community.

\chapter{Forking Process}
\section*{The Heart of Collaboration on GitHub}
The beauty of open-source lies in its democratization of content creation. In this section, we demystify the process of "forking" on GitHub, guiding you step-by-step on how to take CiB and create a version uniquely yours.

\subsection*{Understanding Forking}
Before diving into the specifics, it's crucial to understand what "forking" means in the context of GitHub. In the simplest terms, to "fork" a project means to create a personal copy of someone else's project. Forking allows you to freely experiment with changes without affecting the original project. Forking is akin to taking a book you admire and making a copy to write your notes, edits, or additional chapters without altering the original book.

\subsection*{Why Fork?}
\subsubsection*{Experimentation}
It provides a safe space where you can test out ideas, make changes, or introduce new content.

\subsubsection*{Personalization}
For projects like CiB, it allows readers to customize the content, tailor it to their perspectives, or even localize it for specific audiences.

\subsubsection*{Collaboration}
If you believe your changes have broad appeal, you can propose that they be incorporated back into the original project, enriching it with your unique contributions.

\subsection*{Step-by-Step Forking Guide}
\subsubsection*{Set Up Your GitHub Account}
If you don't have an account on GitHub, you'll need to create one. Visit GitHub's official site and sign up.

\subsubsection*{Navigate to the CiB Repository}
Once logged in, search for the CiB project or navigate to its URL directly.

\subsubsection*{Click the 'Fork' Button}
The fork button is located at the top right corner of the repository page; this button will create a copy of CiB in your account.

\subsubsection*{Clone Your Forked Repository}
Forking allows you to have a local copy on your computer, making editing and experimentation easier. Use the command: \texttt{git clone [URL of your forked repo]}.

\subsubsection*{Make Your Changes}
Using your preferred tools, introduce the edits, additions, or modifications you desire.

\subsubsection*{Commit and Push Changes}
Once satisfied, save these changes (known as a "commit") and then "push" them to your forked repository on GitHub.

\subsubsection*{Optional – Create a Pull Request}
If you believe your changes should be incorporated into the original CiB repository, you can create a "pull request." A pull request notifies the original authors of your suggestions.

\subsection*{Things to Keep in Mind}
\subsubsection*{Stay Updated}
The original CiB project may undergo updates. It's a good practice to regularly "pull" from the original repo to keep your fork up-to-date.

\subsubsection*{Engage with the Community}
Open-source thrives on community interactions. Engage in discussions, seek feedback, and please remain open to constructive criticism.

\subsection*{Conclusion: Embracing the Forking Culture}
Forking is more than just a technical process; it symbolizes the ethos of open-source — a world where knowledge is not hoarded but shared, refined, and built upon collectively. By forking CiB or any other project, you're not just creating a personal copy; you're becoming a part of a global movement that values collaboration, innovation, and the shared pursuit of knowledge. So, embark on this journey, make your unique mark, and contribute to the ever-evolving corpus of collective wisdom.

\chapter{Editing and Customizing}
\section*{Tailoring Repositories to Suit Your Needs}
Now, let's build upon the forking process; this segment delves into the next steps. How can you edit and customize your version of CiB? What tools and techniques are available at your disposal? Embark on this informative journey as we guide you through the intricacies of editing on GitHub.

\subsection*{Understanding the GitHub Workspace}
Before diving into the specifics of editing, it's essential to familiarize yourself with the GitHub workspace. Think of it as a digital toolshed where each tool serves a unique function:

\begin{itemize}
    \item \textbf{Repository (Repo)}: This is the project's main folder where all your project's files are stored and where you track all changes.
    \item \textbf{Branches}: These are parallel versions of a repository, allowing you to work on features or edits without altering the main project.
    \item \textbf{Commits}: This is a saved change in the repository, akin to saving a file after making edits.
    \item \textbf{Pull Requests}: This is how you notify the main project of desired changes, proposing that your edits be merged with the original.
\end{itemize}

\subsection*{Editing Files Directly on GitHub}
For minor changes, you might opt to edit directly on GitHub:

\begin{enumerate}
    \item Navigate to the File: Within your forked CiB repository, find the file you want to edit.
    \item Click the Pencil Icon: This button allows you to edit the file.
    \item Make Your Edits: Modify the content as needed.
    \item Save and Commit: Below the editing pane, you'll see a "commit changes" section. Add a brief note summarizing your changes and click 'Commit.'
\end{enumerate}

\subsection*{Editing Files Locally}
For extensive customization:

\begin{enumerate}
    \item Clone Your Repository: Use a tool like Git to clone (download) your forked repo to your local computer.
    \item Edit Using Your Preferred Tools: This could range from text editors to specialized software, depending on the file type.
    \item Commit and Push: After making your changes, save them (commit) and then upload (push) them to your GitHub repository.
\end{enumerate}

\subsection*{Utilizing Branches for Extensive Customization}
Branches are especially useful for significant overhauls or when working on different versions:

\begin{enumerate}
    \item Create a New Branch: From your main project page, use the branch dropdown to type in a new branch name and create it.
    \item Switch to Your Branch: Ensure you're working in this new parallel environment.
    \item Make and Commit Changes: As you would in the main project.
    \item Merging: Once satisfied with your edits in the branch, you can merge these changes back into the main project or keep them separate as a different version.
\end{enumerate}

\subsection*{Exploring Additional Tools and Extensions}
GitHub's ecosystem is rich with tools and extensions to enhance your editing experience:

\begin{itemize}
    \item \textbf{GitHub Desktop}: An application that simplifies the process of managing your repositories without using command-line tools.
    \item \textbf{Markdown Editors}: Since many GitHub files (like READMEs) are written in Markdown, tools like StackEdit or Dillinger can be invaluable.
    \item \textbf{Extensions for Browsers}: Tools like Octotree can help in navigating repositories more effortlessly.
\end{itemize}

\subsection*{Conclusion: The Art of Tailored Content}
Editing and customizing on GitHub might seem daunting initially, but with practice, it transforms into a manageable workflow. Many people find that the ability to take a project like CiB and mold it into something uniquely theirs is empowering. It's a testament to the open-source community's ethos, where shared knowledge becomes the canvas and our collective edits, the brushstrokes, crafting an ever-evolving masterpiece. As you embark on your customization journey, remember that every edit, no matter how small, contributes to the project potentially in significant ways.

\chapter{Engaging with the Community}
\section*{Joining the Global Conversation}

\subsection*{The Significance of the GitHub Community}
The digital age has bestowed upon us the gift of connectivity. On platforms like GitHub, this connectivity transcends borders, disciplines, and ideologies, culminating in a melting pot of diverse ideas and knowledge. For mathematicians and math enthusiasts, GitHub offers a space not only to store and manage content but also to engage with an audience that is passionate, informed, and eager to contribute.

\subsection*{1. Discussions and Debates}
One of the most enriching aspects of the GitHub community is the plethora of discussions that unfold:

\begin{itemize}
    \item \textbf{Issues}: A core feature of GitHub, "issues" allow users to raise questions, report problems, or propose enhancements. 
    \item \textbf{GitHub Discussions}: A newer feature, Discussions, acts like a community forum. It's an excellent place for extended conversations, brainstorming, and sharing ideas or resources.
\end{itemize}

\subsection*{2. Collaborative Content Creation}
Beyond solitary endeavors, GitHub shines in its collaborative capabilities:

\begin{itemize}
    \item \textbf{Pull Requests}: If you have made an alteration to a math text or added a new perspective, pull requests are the way to propose these changes to the original repository owner. Pull requests foster a collaborative spirit, where content isn't static but continually evolving with community input.
    \item \textbf{Fork and Merge}: As you've learned, forking allows you to create your version of a repository. Engaging with the Community means you can merge changes from others into your fork, blending a mixture of diverse insights.
\end{itemize}

\subsection*{3. Building and Nurturing Networks}
Connections made on GitHub often spill over into lasting professional relationships:

\begin{itemize}
    \item \textbf{Following and Followers}: Like on social media platforms, you can follow contributors whose work resonates with you. Following contributors creates a curated feed of updates and also allows you to be part of a more extensive network.
    \item \textbf{GitHub Stars}: If a particular project or repository impresses you, give it a star! Starring not only bookmarks the project for you but also shows appreciation to the creator.
\end{itemize}

\subsection*{4. Learning and Growing Through Feedback}
The Community's feedback is an invaluable asset:

\begin{itemize}
    \item \textbf{Code Reviews}: Although traditionally for software, text writers can use this feature to receive feedback on their methodologies or approaches, refining their work.
    \item \textbf{Community Insights}: The "insights" tab on a repository provides analytics. For text writers, this can give a sense of which topics garner more attention and interest.
\end{itemize}

\subsection*{5. Participating in Community Events}
GitHub often hosts and sponsors events:

\begin{itemize}
    \item \textbf{Hackathons}: While traditionally for coders, these events can be repurposed for text writer content creation, where participants collaboratively tackle projects or themes.
    \item \textbf{Webinars and Workshops}: These events can range from mastering GitHub's technical side to thematic discussions on math topics.
\end{itemize}

\subsection*{A Project of Collective Wisdom}
Math, in many ways, is a collective endeavor. GitHub can provide a dynamic Community. By engaging with this Community you can become an active participant in the creation of mathematical texts.

\chapter{Functions}
\section{What is a Function?}
\subsection{Definition and Examples}
\begin{definition}
A function is a relation between a set of inputs (domain) and a set of permissible outputs (range) with the property that each input is related to exactly one output.
\end{definition}

\begin{example}
Consider the function \( f(x) = x^2 \), which maps each real number to its square.
\end{example}

\begin{tikzpicture}
\begin{axis}[
    axis lines=middle,
    xlabel=\(x\),
    ylabel=\(f(x)\),
    xmin=-3, xmax=3,
    ymin=-1, ymax=9,
    grid=both,
    grid style={line width=.1pt, draw=gray!10},
    major grid style={line width=.2pt,draw=gray!50},
    minor tick num=1,
    enlargelimits={abs=0.5},
    axis line style={latex-latex},
    ticklabel style={font=\small,fill=white},
    xlabel style={at={(ticklabel* cs:1)},anchor=north west},
    ylabel style={at={(ticklabel* cs:1)},anchor=south west}
]
\addplot[domain=-3:3, samples=100, smooth, thick, blue] {x^2};
\node[anchor=west] at (axis cs: 2,4) {\(f(x) = x^2\)};
\end{axis}
\end{tikzpicture}

\begin{example}
Consider the function \( g(x) = \sin(x) \), which maps real numbers to their sine values.
\end{example}

\begin{tikzpicture}
\begin{axis}[
    axis lines=middle,
    xlabel=\(x\),
    ylabel=\(g(x)\),
    xmin=-2*pi, xmax=2*pi,
    ymin=-1.5, ymax=1.5,
    grid=both,
    grid style={line width=.1pt, draw=gray!10},
    major grid style={line width=.2pt,draw=gray!50},
    minor tick num=1,
    enlargelimits={abs=0.5},
    axis line style={latex-latex},
    ticklabel style={font=\small,fill=white},
    xlabel style={at={(ticklabel* cs:1)},anchor=north west},
    ylabel style={at={(ticklabel* cs:1)},anchor=south west},
    xtick={-6.28318, -3.14159, 0, 3.14159, 6.28318},
    xticklabels={\(-2\pi\), \(-\pi\), 0, \(\pi\), \(2\pi\)}
]
\addplot[domain=-2*pi:2*pi, samples=100, smooth, thick, red] {sin(deg(x))};
\node[anchor=west] at (axis cs: 3,0.5) {\(g(x) = \sin(x)\)};
\end{axis}
\end{tikzpicture}

\begin{example}
Consider the piecewise function \( h(x) \) defined as \( h(x) = \left\{\begin{array}{ll} x^2 & \text{if } x \leq 0 \\ \sqrt{x} & \text{if } x > 0 \end{array}\right. \).
\end{example}

\begin{tikzpicture}
\begin{axis}[
    axis lines=middle,
    xlabel=\(x\),
    ylabel=\(h(x)\),
    xmin=-3, xmax=3,
    ymin=-1, ymax=3,
    grid=both,
    grid style={line width=.1pt, draw=gray!10},
    major grid style={line width=.2pt,draw=gray!50},
    minor tick num=1,
    enlargelimits={abs=0.5},
    axis line style={latex-latex},
    ticklabel style={font=\small,fill=white},
    xlabel style={at={(ticklabel* cs:1)},anchor=north west},
    ylabel style={at={(ticklabel* cs:1)},anchor=south west}
]
\addplot[domain=-3:0, samples=50, smooth, thick, green] {x^2};
\addplot[domain=0:3, samples=50, smooth, thick, green] {sqrt(x)};
\node[anchor=west] at (axis cs: 2,1.5) {\(h(x)\)};
\end{axis}
\end{tikzpicture}

\subsection{Exercises}
\begin{exercise}
Graph the function \( f(x) = 3x - 2 \). Identify its slope and y-intercept.
\end{exercise}

\begin{exercise}
Consider the function \( g(x) = \frac{1}{x} \). For what values of \( x \) is \( g(x) \) undefined?
\end{exercise}

\begin{exercise}
Find the domain and range of the function \( h(x) = \sqrt{x + 4} \).
\end{exercise}

\begin{exercise}
Determine if the function \( f(x) = x^3 - x \) is even, odd, or neither. Justify your answer.
\end{exercise}

\begin{exercise}
Sketch the graph of the piecewise function \( p(x) = \left\{\begin{array}{ll} x^2 & \text{if } x < 0 \\ x + 1 & \text{if } x \geq 0 \end{array}\right. \). Indicate any points of discontinuity.
\end{exercise}

\begin{exercise}
Given the function \( f(x) = 2x^2 - 5x + 3 \), find \( f(-1) \) and \( f(2) \).
\end{exercise}

\begin{exercise}
For the function \( f(x) = |x - 3| \), find the x-coordinate where \( f(x) = 0 \).
\end{exercise}

\begin{exercise}
Create a function \( f(x) \) that has a domain of all real numbers except \( x = 2 \) and a range of \( y \geq 0 \).
\end{exercise}

\begin{exercise}
Determine the intervals on which the function \( f(x) = -x^2 + 4x - 3 \) is increasing and decreasing.
\end{exercise}

\begin{exercise}
For the function \( g(x) = \cos(x) \), evaluate \( g(\pi/2) \) and \( g(\pi) \).
\end{exercise}

% You can add instructions for submitting solutions or further exploration activities here

\section{Solutions to Exercises}

\begin{solution}[1]
To graph \( f(x) = 3x - 2 \), we note that the slope (m) is 3 and the y-intercept (b) is -2. The graph is a straight line with these characteristics.

\begin{tikzpicture}
\begin{axis}[
    axis lines=middle,
    xlabel=\(x\),
    ylabel=\(f(x)\),
    xmin=-3, xmax=3,
    ymin=-5, ymax=7,
    grid=both,
    grid style={line width=.1pt, draw=gray!10},
    major grid style={line width=.2pt,draw=gray!50},
    minor tick num=1,
    enlargelimits={abs=0.5},
    axis line style={latex-latex},
    ticklabel style={font=\small,fill=white},
    xlabel style={at={(ticklabel* cs:1)},anchor=north west},
    ylabel style={at={(ticklabel* cs:1)},anchor=south west}
]
\addplot[domain=-3:3, samples=100, smooth, thick, blue] {3*x - 2};
\end{axis}
\end{tikzpicture}
\end{solution}

\begin{solution}[2]
The function \( g(x) = \frac{1}{x} \) is undefined when \( x = 0 \) because division by zero is not allowed.
\end{solution}

\begin{solution}[3]
For \( h(x) = \sqrt{x + 4} \), the domain is \( x \geq -4 \) (since the expression inside the square root must be non-negative), and the range is \( y \geq 0 \) (since the square root produces non-negative results).
\end{solution}

\begin{solution}[4]
The function \( f(x) = x^3 - x \) is odd because \( f(-x) = (-x)^3 - (-x) = -x^3 + x = -f(x) \).
\end{solution}

\begin{solution}[5]
The piecewise function \( p(x) \) is sketched below. It is discontinuous at \( x = 0 \) since the left-hand limit (0) does not equal the right-hand limit (1).

\begin{tikzpicture}
\begin{axis}[
    axis lines=middle,
    xlabel=\(x\),
    ylabel=\(p(x)\),
    xmin=-3, xmax=3,
    ymin=-1, ymax=4,
    grid=both,
    grid style={line width=.1pt, draw=gray!10},
    major grid style={line width=.2pt,draw=gray!50},
    minor tick num=1,
    enlargelimits={abs=0.5},
    axis line style={latex-latex},
    ticklabel style={font=\small,fill=white},
    xlabel style={at={(ticklabel* cs:1)},anchor=north west},
    ylabel style={at={(ticklabel* cs:1)},anchor=south west}
]
\addplot[domain=-3:0, samples=50, smooth, thick, green, unbounded coords=jump] {x^2};
\addplot[domain=0:3, samples=50, smooth, thick, green] {x + 1};
\end{axis}
\end{tikzpicture}
\end{solution}

\begin{solution}[6]
For \( f(x) = 2x^2 - 5x + 3 \), we find \( f(-1) = 2(-1)^2 - 5(-1) + 3 = 10 \) and \( f(2) = 2(2)^2 - 5(2) + 3 = -1 \).
\end{solution}

\begin{solution}[7]
The function \( f(x) = |x - 3| \) equals 0 when \( x - 3 = 0 \), i.e., \( x = 3 \).
\end{solution}

\begin{solution}[8]
One such function is \( f(x) = \frac{1}{(x-2)^2} \). The denominator ensures that \( x \neq 2 \) (domain), and the square ensures all output values are positive (range).
\end{solution}

\begin{solution}[9]
For \( f(x) = -x^2 + 4x - 3 \), the function is increasing where its derivative \( f'(x) = -2x + 4 \) is positive, i.e., for \( x < 2 \), and decreasing where \( f'(x) \) is negative, i.e., for \( x > 2 \).
\end{solution}

\begin{solution}[10]
For \( g(x) = \cos(x) \), we have \( g(\pi/2) = \cos(\pi/2) = 0 \) and \( g(\pi) = \cos(\pi) = -1 \).
\end{solution}

\subsection{Domain and Range}
\begin{definition}
The domain of a function is the set of all possible inputs, while the range is the set of all possible outputs.
\end{definition}

\begin{example}
For \( f(x) = \sqrt{x} \), the domain is \( x \geq 0 \) and the range is \( y \geq 0 \). The graph of this function is as follows:

\begin{tikzpicture}
\begin{axis}[
    axis lines=middle,
    xlabel=\(x\),
    ylabel=\(f(x)\),
    xmin=0, xmax=4,
    ymin=0, ymax=2,
    samples=100,
    grid=both,
    grid style={line width=.1pt, draw=gray!10},
    major grid style={line width=.2pt,draw=gray!50},
    minor tick num=1,
    enlargelimits={abs=0.5},
    axis line style={latex-latex},
    ticklabel style={font=\small,fill=white},
    xlabel style={at={(ticklabel* cs:1)},anchor=north west},
    ylabel style={at={(ticklabel* cs:1)},anchor=south west}
]
\addplot[domain=0:4, smooth, thick, blue] {sqrt(x)};
\end{axis}
\end{tikzpicture}
\end{example}

\begin{example}
Consider the function \( g(x) = \frac{1}{x} \). Its domain is \( x \neq 0 \) and its range is \( y \neq 0 \). The graph is given by:

\begin{tikzpicture}
\begin{axis}[
    axis lines=middle,
    xlabel=\(x\),
    ylabel=\(g(x)\),
    xmin=-5, xmax=5,
    ymin=-5, ymax=5,
    samples=100,
    restrict y to domain=-5:5,
    grid=both,
    grid style={line width=.1pt, draw=gray!10},
    major grid style={line width=.2pt,draw=gray!50},
    minor tick num=1,
    enlargelimits={abs=0.5},
    axis line style={latex-latex},
    ticklabel style={font=\small,fill=white},
    xlabel style={at={(ticklabel* cs:1)},anchor=north west},
    ylabel style={at={(ticklabel* cs:1)},anchor=south west}
]
\addplot[domain=-5:-0.1, samples=50, smooth, thick, red] {1/x};
\addplot[domain=0.1:5, samples=50, smooth, thick, red] {1/x};
\end{axis}
\end{tikzpicture}
\end{example}

\begin{example}
For the function \( h(x) = x^2 - 4 \), the domain is all real numbers, but the range is \( y \geq -4 \). Its graph is:

\begin{tikzpicture}
\begin{axis}[
    axis lines=middle,
    xlabel=\(x\),
    ylabel=\(h(x)\),
    xmin=-3, xmax=3,
    ymin=-5, ymax=5,
    samples=100,
    grid=both,
    grid style={line width=.1pt, draw=gray!10},
    major grid style={line width=.2pt,draw=gray!50},
    minor tick num=1,
    enlargelimits={abs=0.5},
    axis line style={latex-latex},
    ticklabel style={font=\small,fill=white},
    xlabel style={at={(ticklabel* cs:1)},anchor=north west},
    ylabel style={at={(ticklabel* cs:1)},anchor=south west}
]
\addplot[domain=-3:3, smooth, thick, green] {x^2 - 4};
\end{axis}
\end{tikzpicture}
\end{example}


\subsection{Exercises on Domain and Range}

\begin{exercise}
Determine the domain and range of the function \( f(x) = 2x + 3 \).
\end{exercise}

\begin{exercise}
Find the domain and range of the function \( g(x) = \frac{1}{x - 2} \).
\end{exercise}

\begin{exercise}
Identify the domain and range for the quadratic function \( h(x) = x^2 - 6x + 9 \).
\end{exercise}

\begin{exercise}
For the function \( k(x) = \ln(x - 1) \), determine its domain and range.
\end{exercise}

\begin{exercise}
Consider the function \( m(x) = \frac{x}{x^2 - 4} \). What are its domain and range?
\end{exercise}

\begin{exercise}
Find the domain and range for the trigonometric function \( n(x) = \sin(x) \).
\end{exercise}

\begin{exercise}
Determine the domain and range of the absolute value function \( p(x) = |x + 5| \).
\end{exercise}

\begin{exercise}
For the piecewise function \( q(x) = \begin{cases} x^2 & \text{if } x \leq 0 \\ 2x + 3 & \text{if } x > 0 \end{cases} \), identify the domain and range.
\end{exercise}

\begin{exercise}
Consider the exponential function \( r(x) = 2^x \). What is its domain and range?
\end{exercise}

\begin{exercise}
Identify the domain and range for the cubic function \( s(x) = x^3 - 3x \).
\end{exercise}

\subsection{Solutions to Exercises on Domain and Range}

\begin{solution}[1]
For \( f(x) = 2x + 3 \), the domain is all real numbers since there are no restrictions on \( x \). The range is also all real numbers because as \( x \) takes any real value, \( 2x + 3 \) also covers all real numbers.
\end{solution}

\begin{solution}[2]
The function \( g(x) = \frac{1}{x - 2} \) is undefined when \( x = 2 \), so the domain is \( x \neq 2 \). The range is all real numbers except \( y \neq 0 \) because the function never touches the y-axis.
\end{solution}

\begin{solution}[3]
For the quadratic function \( h(x) = x^2 - 6x + 9 \), the domain is all real numbers. Since the vertex of this parabola is at \( x = 3 \) and it opens upwards, the minimum value is \( h(3) = 0 \), thus the range is \( y \geq 0 \).
\end{solution}

\begin{solution}[4]
The function \( k(x) = \ln(x - 1) \) requires \( x - 1 > 0 \), so the domain is \( x > 1 \). The range of a natural logarithm function is all real numbers.
\end{solution}

\begin{solution}[5]
For \( m(x) = \frac{x}{x^2 - 4} \), the function is undefined for \( x = \pm 2 \), so the domain is \( x \neq \pm 2 \). The range is all real numbers, as the function can take any y-value.
\end{solution}

\begin{solution}[6]
The domain of the trigonometric function \( n(x) = \sin(x) \) is all real numbers. The range of the sine function is between -1 and 1, inclusive, so \( y \in [-1, 1] \).
\end{solution}

\begin{solution}[7]
For the absolute value function \( p(x) = |x + 5| \), the domain is all real numbers. The range is \( y \geq 0 \) since absolute values are always non-negative.
\end{solution}

\begin{solution}[8]
The piecewise function \( q(x) \) is defined for all real numbers, so the domain is all real numbers. To find the range, consider each piece: \( x^2 \) and \( 2x + 3 \). The range is all non-negative values for \( x^2 \) and all real numbers for \( 2x + 3 \), thus the range is all real numbers.
\end{solution}

\begin{solution}[9]
For the exponential function \( r(x) = 2^x \), the domain is all real numbers, as exponents can take any real value. The range is \( y > 0 \) since exponential functions are always positive.
\end{solution}

\begin{solution}[10]
The cubic function \( s(x) = x^3 - 3x \) has a domain of all real numbers. The range is also all real numbers because as a cubic function, it extends infinitely in both the positive and negative y-directions.
\end{solution}


\subsection{Types of Functions}
\subsubsection{Linear Functions}
\begin{definition}
A linear function has the form \( f(x) = mx + b \), where \( m \) is the slope of the line and \( b \) is the y-intercept.
\end{definition}

Linear functions are fundamental in calculus and represent the simplest form of polynomial functions. Their graph is a straight line, and they have constant rates of change.

\paragraph{Properties of Linear Functions:}
\begin{itemize}
    \item The slope \( m \) determines the steepness of the line and its direction (increasing or decreasing).
    \item The y-intercept \( b \) is the point where the line crosses the y-axis.
    \item Linear functions have a constant rate of change, which is equal to the slope \( m \).
\end{itemize}

\paragraph{Graphing a Linear Function:}
Consider the function \( f(x) = 2x + 3 \). Here, \( m = 2 \) and \( b = 3 \). The graph can be plotted as follows:

\begin{tikzpicture}
\begin{axis}[
    axis lines = left,
    xlabel = \( x \),
    ylabel = {\( f(x) \)},
    xmin=-10, xmax=10,
    ymin=-10, ymax=10,
    grid=both,
    grid style={line width=.1pt, draw=gray!10},
    major grid style={line width=.2pt,draw=gray!50},
    minor tick num=1,
    enlargelimits={abs=0.5},
    axis line style={latex-latex},
    ticklabel style={font=\small,fill=white},
    xlabel style={at={(ticklabel* cs:1)},anchor=north west},
    ylabel style={at={(ticklabel* cs:1)},anchor=south west}
]
% Plotting the line
\addplot[domain=-10:10, samples=100, color=blue]{2*x + 3};
\node[anchor=west] at (axis cs: 5,13) {\(f(x) = 2x + 3\)};
\end{axis}
\end{tikzpicture}

This line intersects the y-axis at (0, 3) and has a slope of 2, indicating it rises 2 units for every 1 unit it moves to the right.

\paragraph{Slope-Intercept Form:}
The equation \( f(x) = mx + b \) is known as the slope-intercept form of a linear function, where the slope \( m \) and y-intercept \( b \) are easily identifiable.

\paragraph{Applications of Linear Functions:}
Linear functions model relationships with a constant rate of change and are widely used in various fields such as economics, physics, and social sciences.

\begin{exercise}
Find the slope and y-intercept of the linear function \( g(x) = -3x + 7 \) and sketch its graph.
\end{exercise}

\begin{exercise}
If a linear function passes through the points (1, 2) and (3, 6), determine its equation.
\end{exercise}


\subsubsection*{Exercises on Linear Functions}

\begin{exercise}
Find the slope and y-intercept of the linear function \( g(x) = -3x + 7 \) and sketch its graph.
\end{exercise}

\begin{exercise}
If a linear function passes through the points (1, 2) and (3, 6), determine its equation.
\end{exercise}

\begin{exercise}
Determine the equation of the line that is parallel to \( f(x) = 4x - 5 \) and passes through the point (2, 3).
\end{exercise}

\begin{exercise}
Find the point of intersection of the linear functions \( f(x) = 2x + 1 \) and \( g(x) = -x + 5 \).
\end{exercise}

\begin{exercise}
A rental car company charges a base fee of \$50 and an additional \$20 per day. Write a linear function representing the total cost \( C \) as a function of the number of days \( d \) rented.
\end{exercise}

\begin{exercise}
Graph the linear function \( h(x) = -\frac{1}{2}x + 4 \) and identify where it intersects the x-axis and y-axis.
\end{exercise}

\begin{exercise}
A phone company offers a monthly plan for \$30 with an additional charge of \$0.05 per minute of calls. Write a linear function representing the monthly cost \( M \) as a function of the number of minutes \( m \) used.
\end{exercise}

\begin{exercise}
Determine whether the points (2, 4), (3, 6), and (5, 10) lie on the same linear function. If so, find the equation of the line.
\end{exercise}

\begin{exercise}
Find the slope of the line that passes through the points (-1, -2) and (3, 4).
\end{exercise}

\begin{exercise}
A company’s profit \( P \) in thousands of dollars is linearly related to the number of units \( n \) sold, with a profit of \$4,000 for 500 units and a loss of \$2,000 for 200 units. Determine the linear function that models the profit.
\end{exercise}

\subsubsection*{Solutions to Exercises on Linear Functions}

\begin{solution}[1]
For \( g(x) = -3x + 7 \), the slope (m) is -3, and the y-intercept (b) is 7. The graph is a straight line with a downward slope, crossing the y-axis at (0, 7).
\end{solution}

\begin{solution}[2]
To find the equation of a line passing through (1, 2) and (3, 6), calculate the slope \( m = \frac{6 - 2}{3 - 1} = 2 \). Using the point-slope form: \( y - 2 = 2(x - 1) \), the equation is \( y = 2x \).
\end{solution}

\begin{solution}[3]
A line parallel to \( f(x) = 4x - 5 \) has the same slope, m = 4. Passing through (2, 3), use point-slope form: \( y - 3 = 4(x - 2) \), yielding \( y = 4x - 5 \).
\end{solution}

\begin{solution}[4]
To find the intersection of \( f(x) = 2x + 1 \) and \( g(x) = -x + 5 \), set them equal: \( 2x + 1 = -x + 5 \). Solving gives \( x = \frac{4}{3} \), and substituting back, \( y = \frac{11}{3} \). Thus, they intersect at \( (\frac{4}{3}, \frac{11}{3}) \).
\end{solution}

\begin{solution}[5]
The total cost \( C \) is given by \( C = 50 + 20d \), where \( d \) is the number of days rented.
\end{solution}

\begin{solution}[6]
For \( h(x) = -\frac{1}{2}x + 4 \), the graph intersects the y-axis at (0, 4) and intersects the x-axis at \( x \) where \( h(x) = 0 \). Setting the equation to 0 gives \( x = 8 \). So, it intersects the x-axis at (8, 0).
\end{solution}

\begin{solution}[7]
The monthly cost \( M \) is \( M = 30 + 0.05m \), where \( m \) is the number of minutes used.
\end{solution}

\begin{solution}[8]
The slope between (2, 4) and (3, 6) is 2, and between (3, 6) and (5, 10) is also 2. Since the slope is consistent, they lie on the same line. The line equation, using point-slope form, is \( y = 2x \).
\end{solution}

\begin{solution}[9]
The slope is \( \frac{4 - (-2)}{3 - (-1)} = \frac{6}{4} = \frac{3}{2} \).
\end{solution}

\begin{solution}[10]
Let \( P = mn + b \). Using given points, form equations: \( 4000 = 500m + b \) and \( -2000 = 200m + b \). Solving these simultaneously gives \( m = 20 \) and \( b = -6000 \). So, \( P = 20n - 6000 \).
\end{solution}


% HERE

\subsubsection{Polynomial Functions}
\begin{definition}
A polynomial function is of the form \( f(x) = a_nx^n + a_{n-1}x^{n-1} + \cdots + a_0 \), where \( n \) is a non-negative integer.
\end{definition}

\subsubsection{Rational Functions}
\begin{definition}
A rational function is the ratio of two polynomial functions.
\end{definition}

\subsubsection{Trigonometric Functions}
\begin{definition}
Trigonometric functions include \( \sin(x) \), \( \cos(x) \), and \( \tan(x) \), relating angles to ratios of sides in right triangles.
\end{definition}

\subsubsection{Exponential and Logarithmic Functions}
\begin{definition}
Exponential functions have the form \( f(x) = a^x \) and logarithmic functions are their inverses.
\end{definition}

\section{Function Properties}
\subsection{Continuity}
\begin{definition}
A function is continuous at a point if the limit of the function as it approaches the point is equal to the function value at that point.
\end{definition}

\subsection{Differentiability}
\begin{definition}
A function is differentiable at a point if it has a defined derivative at that point.
\end{definition}

\subsection{Asymptotic Behavior}
\begin{definition}
Asymptotic behavior refers to the behavior of a function as it approaches infinity or a particular value.
\end{definition}

\subsection{Periodicity and Symmetry}
\begin{definition}
A function is periodic if it repeats its values at regular intervals, and it is symmetric if it exhibits mirror symmetry about an axis or point.
\end{definition}

\section{Function Transformations}
\subsection{Translation and Scaling}
\begin{definition}
Translation shifts a function horizontally or vertically, while scaling changes its size.
\end{definition}

\subsection{Reflection and Rotation}
\begin{definition}
Reflection inverts a function across a line, and rotation turns it around a point.
\end{definition}

\chapter{Limits and Continuity}
\section{Introduction to Limits}
\subsection{Definition of a Limit}
\begin{definition}
The limit of \( f(x) \) as \( x \) approaches \( a \) is \( L \) if for every \( \epsilon > 0 \), there exists a \( \delta > 0 \) such that \( |f(x) - L| < \epsilon \) whenever \( 0 < |x - a| < \delta \).
\end{definition}

\subsection{One-sided Limits}
\begin{definition}
The one-sided limits of \( f(x) \) as \( x \) approaches \( a \) from the left (or right) are the values the function approaches as \( x \) approaches \( a \) from the left (or right).
\end{definition}

\subsection{Limits Involving Infinity}
\begin{definition}
A limit involving infinity is the value that \( f(x) \) approaches as \( x \) approaches infinity, or as \( f(x) \) approaches infinity for some finite \( x \).
\end{definition}

\section{Properties of Limits}
\subsection{Limit Laws}
\begin{theorem}
Standard limit laws include the sum law, product law, quotient law, and power law.
\end{theorem}

\subsection{Squeeze Theorem}
\begin{theorem}
If \( f(x) \leq g(x) \leq h(x) \) for all \( x \) near \( a \) and the limits of \( f(x) \) and \( h(x) \) as \( x \) approaches \( a \) are equal, then the limit of \( g(x) \) as \( x \) approaches \( a \) is the same.
\end{theorem}

\subsection{Limits of Trigonometric Functions}
\begin{theorem}
Includes limits such as \( \lim_{x \to 0} \frac{\sin x}{x} = 1 \).
\end{theorem}

\section{Continuity}
\subsection{Definition and Examples}
\begin{definition}
A function is continuous at a point if it is defined at that point, its limit exists at that point, and the limit equals the function value.
\end{definition}

\subsection{Types of Discontinuities}
\begin{definition}
Discontinuities can be classified as removable, jump, or infinite.
\end{definition}

\subsection{Intermediate Value Theorem}
\begin{theorem}
If a function is continuous on a closed interval \([a, b]\) and \( N \) is any number between \( f(a) \) and \( f(b) \), then there exists a number \( c \) in the interval such that \( f(c) = N \).
\end{theorem}

\chapter{The Derivative}
\section{Understanding Derivatives}
\subsection{Definition and Notation}
\begin{definition}
The derivative of a function \( f \) at a point \( a \) is defined as \(\lim_{h \to 0} \frac{f(a+h) - f(a)}{h}\), if this limit exists.
\end{definition}

\subsection{Interpreting Derivatives}
\begin{discussion}
The derivative at a point gives the slope of the tangent line to the function at that point. It represents the rate of change of the function with respect to its variable.
\end{discussion}

\subsection{Derivative as a Function}
\begin{definition}
The derivative of a function \( f \) is itself a function that assigns to each point \( x \) the slope of the tangent to \( f \) at \( x \).
\end{definition}

\section{Techniques of Differentiation}
\subsection{Product Rule and Quotient Rule}
\begin{theorem}
The product rule states that \( (fg)' = f'g + fg' \) and the quotient rule states that \( \left(\frac{f}{g}\right)' = \frac{f'g - fg'}{g^2} \), where \( f \) and \( g \) are functions of \( x \).
\end{theorem}

\subsection{Chain Rule}
\begin{theorem}
If a function \( y \) is a function of \( u \) which is a function of \( x \), then the derivative of \( y \) with respect to \( x \) is given by \( \frac{dy}{dx} = \frac{dy}{du} \cdot \frac{du}{dx} \).
\end{theorem}

\subsection{Higher-Order Derivatives}
\begin{definition}
The \( n \)-th order derivative of \( f \), denoted \( f^{(n)} \), is the derivative of \( f^{(n-1)} \).
\end{definition}

\subsection{Implicit Differentiation}
\begin{method}
Implicit differentiation is used when a function is given in an implicit form \( F(x, y) = 0 \) rather than the explicit form \( y = f(x) \).
\end{method}

\section{Applications of Derivatives}
\subsection{Tangent Lines and Normal Lines}
\begin{application}
The equation of the tangent line to a curve \( y = f(x) \) at a point \( (a, f(a)) \) is \( y - f(a) = f'(a)(x - a) \).
\end{application}

\subsection{Increasing and Decreasing Functions}
\begin{theorem}
A function \( f \) is increasing (decreasing) on an interval if its derivative \( f' \) is positive (negative) on that interval.
\end{theorem}

\subsection{Maxima and Minima}
\begin{method}
Critical points, where \( f'(x) = 0 \) or \( f'(x) \) does not exist, are potential locations of local maxima or minima.
\end{method}

\subsection{Concavity and Inflection Points}
\begin{discussion}
A function is concave up (down) where its second derivative \( f'' \) is positive (negative). Points where \( f'' \) changes sign are inflection points.
\end{discussion}

\subsection{Optimization Problems}
\begin{application}
Optimization involves finding the maximum or minimum values of a function subject to certain constraints, often using derivatives.
\end{application}

\subsection{Related Rates}
\begin{method}
Related rates problems involve finding the rate at which one quantity changes with respect to another using the chain rule.
\end{method}

\chapter{Integration}
\section{Antiderivatives}
\subsection{Definition and Basic Techniques}
\begin{definition}
An antiderivative of a function \( f(x) \) is a function \( F(x) \) such that \( F'(x) = f(x) \).
\end{definition}
\begin{example}
An antiderivative of \( f(x) = 2x \) is \( F(x) = x^2 + C \), where \( C \) is a constant.
\end{example}

\subsection{Indefinite Integrals}
\begin{definition}
The indefinite integral of \( f(x) \), denoted \( \int f(x) \, dx \), represents the collection of all its antiderivatives.
\end{definition}

\subsection{Initial Value Problems}
\begin{problem}
Given a differential equation \( f'(x) = g(x) \) and an initial condition \( f(x_0) = y_0 \), find the function \( f(x) \).
\end{problem}

\section{Definite Integrals}
\subsection{Definition and Interpretation}
\begin{definition}
The definite integral of \( f(x) \) from \( a \) to \( b \), denoted \( \int_a^b f(x) \, dx \), is the limit of the sum of areas of rectangles under the curve of \( f(x) \) as the width of the rectangles approaches zero.
\end{definition}

\subsection{Properties of Definite Integrals}
\begin{theorem}
Properties include linearity, additivity, and the fact that the integral from \( a \) to \( a \) is zero.
\end{theorem}

\subsection{Fundamental Theorem of Calculus}
\begin{theorem}
If \( F \) is an antiderivative of \( f \) on an interval, then \( \int_a^b f(x) \, dx = F(b) - F(a) \).
\end{theorem}

\subsection{Numerical Integration Methods}
\begin{method}
Methods include the Trapezoidal Rule and Simpson's Rule, which provide approximations of definite integrals.
\end{method}

\section{Applications of Integrals}
\subsection{Area Between Curves}
\begin{application}
The area between curves \( f(x) \) and \( g(x) \) from \( a \) to \( b \) is \( \int_a^b |f(x) - g(x)| \, dx \).
\end{application}

\subsection{Volumes of Solids of Revolution}
\begin{application}
The volume of a solid obtained by rotating a region about an axis can be found using the Disk or Washer methods.
\end{application}

\subsection{Arc Length and Surface Area}
\begin{method}
Formulas for the arc length of a curve and the surface area of a solid of revolution involve definite integrals.
\end{method}

\subsection{Work and Fluid Forces}
\begin{application}
Work done in moving an object and the force exerted by a fluid can be calculated using integration.
\end{application}

\chapter{Advanced Topics in Calculus}
\section{Sequences and Series}
\subsection{Convergence and Divergence}
\begin{definition}
A sequence \( \{a_n\} \) converges if it approaches a limit as \( n \) goes to infinity. A series \( \sum a_n \) converges if the sequence of its partial sums converges.
\end{definition}

\subsection{Tests for Convergence}
\begin{method}
Tests include the Integral Test, Comparison Test, Ratio Test, and Root Test.
\end{method}

\subsection{Power Series and Taylor Series}
\begin{definition}
A power series is a series of the form \( \sum_{n=0}^\infty a_n (x - c)^n \). A Taylor series is a power series that represents a function.
\end{definition}

\section{Multivariable Calculus}
\subsection{Functions of Several Variables}
\begin{definition}
A function of several variables is a function that takes several inputs and produces a single output.
\end{definition}

\subsection{Partial Derivatives}
\begin{definition}
The partial derivative of a function with respect to one of its variables is its derivative with respect to that variable, holding the other variables constant.
\end{definition}

\subsection{Multiple Integrals}
\begin{definition}
Multiple integrals involve integration over more than one variable, such as double and triple integrals.
\end{definition}

\subsection{Vector Calculus}
\begin{discussion}
Vector calculus involves differentiation and integration of vector fields, including topics like gradient, divergence, and curl.
\end{discussion}

\section{Differential Equations}
\subsection{First Order Differential Equations}
\begin{method}
Methods for solving include separation of variables, integrating factors, and exact equations.
\end{method}

\subsection{Second Order Linear Differential Equations}
\begin{method}
Includes homogeneous and non-homogeneous equations, characteristic equations, and particular solutions.
\end{method}

\subsection{Systems of Differential Equations}
\begin{discussion}
Systems of differential equations involve several interrelated differential equations and can be solved using matrix methods and eigenvalues.
\end{discussion}

\section{Conclusion}
\paragraph{}
This concludes the math currently included in CiB. Please fork the LaTeX source code for CiB (available on GitHub) and create your own book that chooses the facts and exercises most relevant to you! Also, starring the CiB project on GitHub would be greatly appreciated! Thanks for reading CiB!


% --- Appendices ---
\clearpage
\addcontentsline{toc}{chapter}{Appendices}
\appendix
\renewcommand{\thechapter}{\Roman{chapter}} % Ensuring chapters are numbered as I, II, III, etc.

%\appendix
\chapter{Basic GitHub Guide}
\section*{A Quick Start to Your GitHub Journey}

Welcome to the fascinating world of GitHub, a platform that has revolutionized the way we collaborate on projects, share code, and build software together. Whether you are a programmer, a writer, or a mathematician, GitHub provides a set of powerful tools to help you collaborate with others, manage your projects, and contribute to the vast world of open-source software. In this guide, we will walk you through the foundational steps to get started with GitHub, helping you to navigate, contribute, and make the most out of this incredible platform.

\subsection*{Creating Your GitHub Account}

The first step to joining the GitHub community is to create an account. Here’s how you can do it:

\begin{enumerate}
    \item Visit the \href{https://github.com/}{GitHub website}.
    \item Click on the “Sign up” button.
    \item Fill in the required information, including your username, email address, and password.
    \item Verify your account and complete the sign-up process.
\end{enumerate}

Once you have created your account, take a moment to explore your new GitHub dashboard. Here, you will find a variety of tools and features that will help you manage your projects, collaborate with others, and discover new and interesting repositories.

\subsection*{Creating Your First Repository}

A repository (or “repo”) is a digital directory where you can store your project files. Here’s how you can create your first repository:

\begin{enumerate}
    \item From your GitHub dashboard, click on the “New” button to create a new repository.
    \item Give your repository a name and provide a brief description.
    \item Initialize this repository with a README file. (This is an optional step, but it’s a good practice to include a README file in every repository to explain what your project is about.)
    \item Click “Create repository.”
\end{enumerate}

Congratulations! You have just created your first GitHub repository. You can now start adding files, collaborating with others, and managing your project right from GitHub.

\subsection*{Making Changes and Commits}

GitHub uses Git, a version control system, to keep track of changes made to your project. Here’s a quick guide on how to make changes and commits:

\begin{enumerate}
    \item Navigate to your repository on GitHub.
    \item Find the file you want to edit, and click on it.
    \item Click the pencil icon to start editing.
    \item Make your changes and then scroll down to the “Commit changes” section.
    \item Provide a commit message that explains the changes you made.
    \item Choose whether you want to commit directly to the main branch or create a new branch for your changes.
    \item Click “Commit changes.”
\end{enumerate}

Your changes are now saved, and a new commit is created. Every commit has a unique ID, making it easy to track changes, revert to previous versions, and collaborate with others.

\subsection*{Collaborating with Others}

One of the biggest strengths of GitHub is its collaborative nature. Here are some ways you can collaborate with others:

\begin{itemize}
    \item \textbf{Forking:} You can fork a repository, create your own copy, make changes, and then propose those changes back to the original project.
    \item \textbf{Issues:} Use issues to report bugs, request new features, or start a discussion with the community.
    \item \textbf{Pull Requests:} Propose changes to a project by creating a pull request. This allows others to review your changes, discuss them, and eventually merge them into the project.
\end{itemize}

\subsection*{Conclusion: Embarking on Your GitHub Adventure}

Now that you have a basic understanding of GitHub and how it works, you are ready to embark on your GitHub adventure. Explore repositories, contribute to open-source projects, collaborate with others, and build amazing things together. Remember, the GitHub community is vast and supportive, and there is a wealth of knowledge and resources available to help you along the way. Happy coding!


\chapter{Basic Python and Colab Guide}
\section*{Introduction to Python for Calculus}

Python's versatility in mathematics, science, engineering, and data analysis stems from its simplicity and extensive library ecosystem. This guide will delve into Python packages essential for math and calculus, showcasing their utility in Google Colab notebooks.

\subsection*{Setting Up Python and Google Colab}

Google Colab is a free, cloud-based platform enabling Python programming in a browser. It offers free computational resources, ideal for Python learning and experimentation.

Visit \href{https://colab.research.google.com/}{Google Colab} to start. Create a new notebook, and use code cells to write and execute Python code with Shift+Enter.

\subsection*{Important Python Packages for Calculus}

\subsubsection*{NumPy}
NumPy, fundamental for scientific computing, offers support for large, multi-dimensional arrays and matrices, along with various mathematical functions.

\subsubsection*{SymPy}
SymPy, a library for symbolic mathematics, allows algebraic manipulations and equation solving symbolically, perfect for calculus operations like differentiation and integration.

\subsubsection*{Matplotlib}
Matplotlib, a Python plotting library, creates static, interactive, and animated visualizations, useful for graphing functions and data in calculus.

\subsubsection*{Pandas}
Pandas provide high-performance, easy-to-use data structures, and data analysis tools, invaluable for handling and analyzing mathematical data.

\subsubsection*{SciPy}
SciPy extends NumPy by adding a collection of algorithms and high-level commands for data manipulation and visualization.

\subsection*{Examples and Applications}

\subsubsection*{Calculating Derivatives and Integrals with SymPy}
Illustrate using SymPy to compute derivatives and integrals of functions.

\subsubsection*{Data Visualization with Matplotlib and Pandas}
Demonstrate graphing functions and datasets, highlighting calculus concepts.

\subsubsection*{Solving Equations with SciPy}
Show how to solve equations that commonly appear in calculus.

\subsubsection*{Numerical Methods in Python}
Discuss Python's capabilities in numerical differentiation and integration, useful in calculus.

\subsubsection*{Using Python for Limits and Continuity}
Examples showcasing how Python can help in understanding limits and continuity in functions.

\subsection*{Interactive Learning with Google Colab}
Highlight the benefits of using Colab notebooks for interactive calculus learning, including real-time collaboration and easy sharing.


\subsection*{Creating a Colab Notebook for Practice Problems}

In this section, we will guide you through creating a Google Colab notebook to solve calculus practice problems using Python.

\subsubsection*{Setting Up Your Colab Notebook}
To start solving calculus problems with Python:

\begin{enumerate}
    \item Open \href{https://colab.research.google.com/}{Google Colab}.
    \item Choose 'New Notebook' to create a blank notebook.
    \item Rename the notebook to something descriptive, like 'Calculus Practice'.
\end{enumerate}

\subsubsection*{Installing Necessary Libraries}
At the beginning of your notebook, install any necessary Python libraries. For these exercises, ensure NumPy, SymPy, and Matplotlib are available:

\begin{verbatim}
Remove # if the following packages are not installed:
# !pip install numpy sympy matplotlib
\end{verbatim}

\subsubsection*{Solving Exercise 1: Graphing a Linear Function}
Let's solve the first exercise, which involves graphing a linear function.

\begin{verbatim}
import numpy as np
import matplotlib.pyplot as plt

# Define the function
def f(x):
    return 3*x - 2

# Generate x values
x = np.linspace(-10, 10, 400)

# Plot the function
plt.plot(x, f(x))
plt.xlabel('x')
plt.ylabel('f(x)')
plt.title('Graph of f(x) = 3x - 2')
plt.grid(True)
plt.show()

# Slope and y-intercept
print("Slope: 3")
print("Y-intercept: -2")
\end{verbatim}

\subsubsection*{Solving Exercise 2: Identifying Undefined Points in a Function}
Now, let's address the second exercise, which requires identifying for what values the function \( g(x) = \frac{1}{x} \) is undefined.

\begin{verbatim}
import sympy as sp

x = sp.symbols('x')
g = 1/x

# Display the function
sp.plot(g, (x, -10, 10), title='Graph of g(x) = 1/x',
        xlabel='x', ylabel='g(x)', ylim=(-10, 10))

# Identifying undefined points
print("g(x) is undefined for x = 0")
\end{verbatim}

\subsection*{Accessing the Completed Colab Notebook}

The Colab notebook we've discussed is available for viewing and interaction. You can access it by clicking on the following link: \href{https://colab.research.google.com/drive/1HF-cmwqfIZ803i1i-CFyR7ssWsDt7WvV}{Finished Colab Notebook on Graphing and Analysis}. This link will take you directly to the notebook hosted on Google Colab, where you can view the complete code and run it yourself.

\url{https://colab.research.google.com/drive/1HF-cmwqfIZ803i1i-CFyR7ssWsDt7WvV}

\subsection*{Adding the Notebook to Your GitHub Repository}

If you have downloaded the Colab notebook to your local machine and want to add it to your Git repository, follow these terminal commands on your Ubuntu machine:

\begin{verbatim}
# Navigate to your local Git repository directory
cd path/to/your/repo

# Add the Colab notebook file to the repository
git add name_of_the_notebook.ipynb

# Commit the addition with a descriptive message
git commit -m "Add Colab notebook for calculus exercises"

# Push the changes to the remote GitHub repository
git push origin main
\end{verbatim}

%Replace `path/to/your/repo` with the actual path to your local Git repository and `name_of_the_notebook.ipynb` with the filename of your Colab notebook.

\subsubsection*{Using Colab Notebooks for Problem Solving}
These examples demonstrate how you can use Google Colab and Python to solve and visualize calculus problems. You can use similar steps to tackle other exercises, explore different functions, and deepen your understanding of calculus concepts.

\subsection*{Conclusion: Interactive Learning with Colab Notebooks}

Google Colab notebooks offer an interactive and accessible way to explore calculus using Python. By integrating theoretical concepts with computational examples, students can gain a deeper understanding of calculus. We encourage you to use these notebooks to solve exercises, visualize mathematical concepts, and explore the vast possibilities that Python and Colab offer.


\subsection*{Conclusion: Python and Colab in Calculus}

Python, with its comprehensive libraries, offers a powerful toolset for calculus exploration. Combined with Google Colab, it provides an accessible, interactive platform for learning and experimentation. Embrace Python and Colab to enhance your understanding of calculus and to explore mathematical problems creatively and efficiently.


\chapter{Basic \LaTeX\ Guide}
\section*{A Quick Start to Your \LaTeX\ Journey}

Welcome to the immersive world of \LaTeX, a typesetting system widely used for creating scientific and professional documents due to its powerful handling of formulas and bibliographies. This guide is designed to offer you the foundational steps to grasp the basics of \LaTeX, enabling you to craft documents of high typographic quality akin to this book.

\subsection*{Setting Up Your \LaTeX\ Environment}

Before you can start creating documents with \LaTeX, you need to set up a working \LaTeX\ environment on your computer. Here's how you can do it:

\begin{enumerate}
    \item Download and install a \TeX\ distribution, which includes \LaTeX. For Windows, MiKTeX is a popular choice, while Mac users might prefer MacTeX, and TeX Live is widely used on Linux.
    \item Install a \LaTeX\ editor. Some popular options include TeXShop (for Mac), TeXworks (cross-platform), and Overleaf (an online \LaTeX\ editor).
    \item Ensure that your \TeX\ distribution and \LaTeX\ editor are properly configured and integrated.
\end{enumerate}

\subsection*{Creating Your First \LaTeX\ Document}

Once your \LaTeX\ environment is set up, you are ready to create your first \LaTeX\ document. Follow these steps:

\begin{enumerate}
    \item Open your \LaTeX\ editor and create a new document.
    \item Insert the following code to set up a basic \LaTeX\ document:

\begin{verbatim}
\documentclass{article}
\begin{document}
Hello, \LaTeX\ world!
\end{document}
\end{verbatim}

    \item Save your document with a .tex file extension.
    \item Compile your document using your \LaTeX\ editor. This process converts your .tex file into a PDF document.
    \item View the output PDF and admire your first \LaTeX\ creation.
\end{enumerate}

\subsection*{Understanding \LaTeX\ Commands and Environments}

\LaTeX\ documents are created using a series of commands and environments. Commands typically start with a backslash \textbackslash\ and are used to format text, insert special characters, or execute functions. Environments are used to define specific sections of your document that require special formatting.

\begin{itemize}
    \item \textbf{Commands:} For example, \textbackslash\textit\{italics\} will render the word "italics" in italic font.
    \item \textbf{Environments:} To create a bulleted list, you would use the \textit{itemize} environment:

\begin{verbatim}
\begin{itemize}
    \item First item
    \item Second item
\end{itemize}
\end{verbatim}
\end{itemize}

\subsection*{Adding Structure to Your Document}

\LaTeX\ makes it easy to structure your documents with sections, subsections, and chapters. Here’s how you can add structure:

\begin{verbatim}
\section{Introduction}
This is the introduction of your document.
\subsection{Background}
This subsection provides background information.
\subsubsection{Details}
This is a subsubsection for more detailed information.
\end{verbatim}

\subsection*{Including Mathematical Formulas}

\LaTeX\ excels at typesetting mathematical formulas. Use the \textit{equation} environment or the \textdollar\ sign for inline formulas. For example:

\begin{verbatim}
The quadratic formula is \( x = \frac{-b \pm \sqrt{b^2 - 4ac}}{2a} \).
\end{verbatim}

\subsection*{Adding Images and Tables}

You can also include images and tables in your \LaTeX\ documents:

\begin{itemize}
    \item \textbf{Images:} Use the \textit{graphicx} package and the \textit{includegraphics} command.
    \item \textbf{Tables:} Use the \textit{tabular} environment to create tables.
\end{itemize}

\subsection*{Compiling Your Document}

\LaTeX\ documents need to be compiled to produce a PDF. This can be done through your \LaTeX\ editor. If your document includes bibliographies or cross-references, you may need to compile multiple times.

\subsection*{Conclusion: Embracing the Power of \LaTeX}

Congratulations! You have taken your first steps into the world of \LaTeX. With practice, you will discover that \LaTeX\ is a powerful tool for creating professional-quality documents, from simple articles to complex books. Embrace the learning curve, explore the vast array of packages available, and join the community of \LaTeX\ users who are ready to help you on your journey. Happy typesetting!

% --- Bibliography ---
\addcontentsline{toc}{chapter}{Bibliography}
\bibliographystyle{alpha}
\bibliography{references} % Assuming you have a references.bib file

% --- Index ---
% \addcontentsline{toc}{chapter}{Index}
% \printindex

\end{document}
